---
title: "Assignment 4"
layout: page
---
# 1

```{r warning=FALSE, message=FALSE}

pkgs <- c("rvest", "tidyverse", "stringr")
invisible(lapply(pkgs, function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}))

library(rvest)
library(tidyverse)
library(stringr)

# URL
url <- "https://en.wikipedia.org/wiki/List_of_countries_by_foreign-exchange_reserves"

page <- read_html(url)

raw_table <- page %>%
  html_nodes("table.wikitable") %>%
  html_table(fill = TRUE)


df <- raw_table[[1]]

df <- df[3:198, c(1,2,3,5,7)]

colnames(df) <- c("Country", "Continent", "ForEx_Inc_Gold", "ForEx_Exc_Gold", "ReportDate")

df_clean <- df %>%
  mutate(across(everything(), ~ str_remove_all(.x, "\\[\\d+\\]"))) %>%
  mutate(across(everything(), ~ trimws(.x))) %>%   
  filter(!is.na(Country), Country != "") %>%
  mutate(
    ForEx_Inc_Gold = as.numeric(str_remove_all(ForEx_Inc_Gold, ",")),
    ForEx_Exc_Gold = as.numeric(str_remove_all(ForEx_Exc_Gold, ","))
  ) %>%
  select(Country, Continent, ForEx_Inc_Gold, ForEx_Exc_Gold, ReportDate)

df_clean$ReportDate <- as.Date(df_clean$ReportDate, format = "%d %b %Y")
#result
head(df_clean)

```

# Scrapping Other Table

```{r warning=FALSE, message=FALSE}

# URL
url <- "https://en.wikipedia.org/wiki/List_of_Asian_countries_by_population"


page <- read_html(url)

raw_table <- page %>%
  html_nodes("table.wikitable") %>%
  html_table(fill = TRUE)

df <- raw_table[[1]]

df <- df[, c(2,3,5,6,8)]

colnames(df) <- c("Country", "%Asia", "Population", "%Growth", "ReportDate")

df_clean <- df %>%
  mutate(across(everything(), ~ str_remove_all(.x, "\\[\\d+\\]"))) %>%
  mutate(across(everything(), ~ trimws(.x))) %>%   
  filter(!is.na(Country), Country != "") %>%
  mutate(
    Population = as.numeric(str_remove_all(Population, ",")),
  ) %>%
  select(Country, `%Asia`, Population, `%Growth`, ReportDate)

df_clean$ReportDate <- as.Date(df_clean$ReportDate, format = "%d %b %Y")
#result
head(df_clean)

```

# Discussion

I modified the code to scrape the Wikipedia page listing Aian countries by population. Using a similar approach, virtually any HTML-based table can be extracted and transformed into a tidy data frame. The script leverages the rvest package to read the page’s HTML, then selects the target table using XPath. After extraction, I cleaned the data by stripping footnote markers and citations from the country and date columns, and converted the population and world-share percentage columns to numeric format to facilitate analysis. To confirm the process worked correctly, I displayed the first 10 rows of the cleaned data frame.
