---
title: "Assignment 3"
layout: page
---

# Murrell01.R 

```{r}
# Scatterplot with modified elements

x_vals <- c(1, 3, 6, 9, 11, 15) # New x-values for the plot
y1_vals <- c(0.8, 1.2, 1.7, 2.8, 3.4, 4.5) # New y1 data
y2_vals <- c(3.5, 1.0, 0.7, 0.5, 0.3, 0.2) # New y2 data


plot.new()
plot.window(range(x_vals), c(0, 5)) 
lines(x_vals, y1_vals)
lines(x_vals, y2_vals) 
points(x_vals, y1_vals, pch=17, cex=2.5) 
points(x_vals, y2_vals, pch=22, bg="lightblue", cex=2.5) 
par(col="darkgray", fg="darkgray", col.axis="darkgray") 
axis(1, at=seq(0, 16, 3)) 
axis(2, at=seq(0, 5, 1)) 
axis(4, at=seq(0, 5, 1)) 
box(bty="l") 
mtext("Time (seconds)", side=1, line=2, cex=0.8) # Label for x-axis
mtext("Frequency per Travel", side=2, line=2, las=0, cex=0.8) # Label for y1-axis
mtext("Frequency per Second", side=4, line=2, las=0, cex=0.8) # Label for y2-axis
text(6, 4, "Bird 151", col="red") 


```


# Anscombe01.R

```{r}
data(anscombe)

lm1 <- lm(y1 ~ x1, data=anscombe)
lm2 <- lm(y2 ~ x2, data=anscombe)
lm3 <- lm(y3 ~ x3, data=anscombe)
lm4 <- lm(y4 ~ x4, data=anscombe)

op <- par(mfrow = c(2, 2), mar = 0.1 + c(4, 4, 1, 1), oma = c(0, 0, 2, 0))

ff <- y ~ x
mods <- list(lm1, lm2, lm3, lm4)

for (i in 1:4) {
  ff[2:3] <- lapply(paste0(c("y", "x"), i), as.name)
  plot(ff, data = anscombe, col = "darkgreen", pch = 19, bg = "lightgreen", cex = 1.5,
       xlim = c(2, 20), ylim = c(2, 14))
  abline(coef(mods[[i]]), col = "purple", lwd = 2)
}

mtext("Anscombe's 4 Regression Data Sets", outer = TRUE, cex = 1.5)

# Reset plotting parameters
par(op)
```

## Analysis of Anscombe's Data Sets

The four regression plots illustrate the relationship between xx and yy. Despite having identical slopes and coefficients, the true functional relationships vary significantly across the datasets.

Each plot includes a best-fit line that minimizes the squared distance between the data points and the regression line. However, some relationships are better captured by a quadratic function rather than the conventional linear model.

The plots also demonstrate the impact of outlier points. In one dataset, an outlier substantially skews the apparent relationship, emphasizing the need to account for such anomalies in regression analysis.

In another dataset, the xx variable (or yy) exhibits minimal variation, yet a linear relationship still emerges due to the alignment of the remaining points.

# Alternative ways to plot

```{r}
data(anscombe)

lm1 <- lm(y1 ~ x1, data = anscombe)
lm2 <- lm(y2 ~ x2, data = anscombe)
lm3 <- lm(y3 ~ x3, data = anscombe)
lm4 <- lm(y4 ~ x4, data = anscombe)

mods <- list(lm1, lm2, lm3, lm4)

op <- par(mfrow = c(2, 2), mar = 0.1 + c(4, 4, 1, 1), oma = c(0, 0, 2, 0))

for (i in 1:4) {
  plot(anscombe[[paste0("x", i)]], anscombe[[paste0("y", i)]], 
       col = "blue", pch = 1, bg = "gray30", cex = 1.2,
       xlim = c(3, 19), ylim = c(3, 13),
       xlab = "x", ylab = "y", main = paste("Dataset", i))
  abline(mods[[i]], col = "red")
}

mtext("Anscombe's 4 Regression Data Sets", outer = TRUE, cex = 1.5)
par(op)
```


# Fine Tuning

## R basic packages
```{r}
x <- c(0.5, 2, 4, 8, 12, 16)
y1 <- c(1, 1.3, 1.9, 3.4, 3.9, 4.8)
y2 <- c(4, .8, .5, .45, .4, .3)

plot.new()
plot.window(range(x), c(0, 6))

col_y1 <- "darkblue"
col_y2 <- "darkgreen"

lines(x, y1, col = col_y1, lwd = 2)
lines(x, y2, col = col_y2, lwd = 2)

points(x, y1, pch = 17, col = col_y1, cex = 2)
points(x, y2, pch = 15, col = col_y2, cex = 2)

par(col = "gray50", fg = "gray50", col.axis = "gray50", family = "serif")

axis(1, at = seq(0, 16, 4))
axis(2, at = seq(0, 6, 2))
axis(4, at = seq(0, 6, 2))
box(bty = "u")

mtext("Travel Time (s)", side = 1, line = 2, cex = 0.8)
mtext("Responses per Travel", side = 2, line = 2, las = 0, cex = 0.8)
mtext("Responses per Second", side = 4, line = 2, las = 0, cex = 0.8)

text(4, 5, "Bird 131", col = "black", family = "serif")


```
## with tidyverse 

```{r}
library(tidyverse)

# Data points for x, y1, y2
data <- tibble(
  x = c(0.5, 2, 4, 8, 12, 16),
  y1 = c(1, 1.3, 1.9, 3.4, 3.9, 4.8),
  y2 = c(4, 0.8, 0.5, 0.45, 0.4, 0.3)
)

# Plot with ggplot2
ggplot(data) +
  geom_line(aes(x = x, y = y1), color = "darkblue", size = 1.2) +
  geom_line(aes(x = x, y = y2), color = "darkgreen", size = 1.2) +
  geom_point(aes(x = x, y = y1), shape = 17, color = "darkblue", size = 4) +
  geom_point(aes(x = x, y = y2), shape = 15, color = "darkgreen", size = 4) +
  labs(x = "Travel Time (s)", y = "Responses per Travel") +
  annotate("text", x = 4, y = 5, label = "Bird 131", family = "serif", size = 5) +
  theme_minimal(base_family = "serif") +
  theme(
    axis.title = element_text(size = 12, color = "gray50"),
    axis.text = element_text(color = "gray50"),
    axis.line = element_line(color = "gray50"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray80")
  )

```

